Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-10-23 thesis/paper_draft.qmd
refine other cov ests, consistent g(.) for cov only

$$
\hat{\boldsymbol{W}}^{S}_{1} = \lambda_S \, \text{diag}( \hat{\boldsymbol{W}}_1 ) + (1 - \lambda_S) \hat{\boldsymbol{W}}_1 \, ,
$$

where $\text{diag}( \hat{\boldsymbol{W}}_1 )$ is the diagonal matrix comprising the diagonal elements of $\hat{\boldsymbol{W}}_1$. We refer to any $\lambda_S \in [0,1]$ as the shrinkage intensity of the shrinkage estimator. This approach shrinks the covariance matrix $\hat{\boldsymbol{W}}_1$ towards the diagonal target $\text{diag}( \hat{\boldsymbol{W}}_1 )$, meaning the off-diagonal elements are shrunk towards zero while the diagonal ones remain unchanged.

@@

### Problem 2: Latent factors {.unnumbered}

@@

### Problem 3: Proportional scaling {.unnumbered}

@@

To exploit latent common components explicitly, the PC-adjusted method takes the latent factors directly into its construction. It starts by decomposing the covariance matrix $\hat{\boldsymbol{W}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{W}}^K_{1}$ (the correlation matrix after removing the first $K$ principal components). Then we can apply either shrinkage or NOVELIST estimator to $\hat{\boldsymbol{W}}^K_{1}$:

$$
\hat{\boldsymbol{W}}^{g, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + g(\hat{\boldsymbol{W}}^K_{1})
$$

where $g(.)$ is either the shrinkage or NOVELIST estimator, $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$-th largest eigenvalue and the corresponding eigenvector of the sample covariance matrix, respectively. The number of principal components $K$ can be selected using information criteria.

Similar to the NOVELIST estimator, its PC-adjusted variant $\hat{\boldsymbol{W}}^{N, K}_{1}$ requires a cross-validation procedure to select the threshold paramter and adjustment to obtain positive definiteness.

@@

To address the potential issue of the proportionality relationship $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$ not holding in practice. A simple relaxation retains the $1$-step-ahead correlation shape but allows the variances to scale differently with horizon. The scaled variance estimator is given by:

$$
\hat{\boldsymbol{W}}^{g, sv}_{h} = \boldsymbol{D}^{1/2}_h g(\hat{\boldsymbol{R}}_1) \boldsymbol{D}^{1/2}_h,
$$

where $\boldsymbol{D}_h = \text{diag}(\hat{\sigma}^2_{1,h}, \ldots, \hat{\sigma}^2_{n,h})$, and $\hat{\sigma}^2_{i,h}$ is the variance of the $i$-th series' h-step-ahead base forecast errors. Similarly, $g(.)$ is either the shrinkage or NOVELIST estimator.

When NOVELIST is used to produce $h$-step-ahead reconciled forecasts, the cross-validation procedure is slightly modified to evaluate the out-of-sample reconciled forecast MSE using $\hat{\boldsymbol{W}}^{N, sv}_{h}$ instead of $\hat{\boldsymbol{W}}^{N}_{1}$.

@@

Another alternative is to directly estimate the covariance matrix from the h-step-ahead base forecast errors, without assuming any proportionality relationship:

$$
\hat{\boldsymbol{W}}^{g}_{h} = g(\hat{\boldsymbol{W}}_h)
$$

where $\hat{\boldsymbol{W}}_h$ is the covariance matrix of in-sample h-step-ahead base forecast residuals, and $g(.)$ is either the shrinkage or NOVELIST estimator.

Similar to the scaled variance approach, if NOVELIST is used, the cross-validation procedure is modified to compute $\hat{\boldsymbol{W}}^{N}_{h}$ and evaluate the out-of-sample reconciled forecast MSE using it.

>>>>

$$
\hat{\boldsymbol{W}}^{S}_{1} = \lambda_S \, \boldsymbol{D}_1 + (1 - \lambda_S) \hat{\boldsymbol{W}}_1 \, ,
$$

where $\boldsymbol{D}_1 = \text{diag}( \hat{\boldsymbol{W}}_1 )$ is the diagonal matrix comprising the diagonal elements of $\hat{\boldsymbol{W}}_1$ (i.e., the sample variances). We refer to any $\lambda_S \in [0,1]$ as the shrinkage intensity of the shrinkage estimator. This approach shrinks the covariance matrix $\hat{\boldsymbol{W}}_1$ towards the diagonal target $\text{diag}( \hat{\boldsymbol{W}}_1 )$, meaning the off-diagonal elements are shrunk towards zero while the diagonal ones remain unchanged.

@@

### Problem 2: Latent factors {#prob2 .unnumbered}

@@

### Problem 3: Proportional scaling {#prob3 .unnumbered}

@@

To address Problem [2](#prob2) and exploit latent common components explicitly, the PC-adjusted (Principal-Component-adjusted) method takes the latent factors directly into its construction. It starts by decomposing the covariance matrix $\hat{\boldsymbol{W}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\Upsilon}$ (the covariance matrix after removing the first $K$ principal components). Then we can apply either shrinkage or NOVELIST estimator to $\hat{\boldsymbol{W}}^K_{1}$:

$$
\hat{\boldsymbol{W}}^{g, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + g(\hat{\Upsilon})
$$

where $g(.)$ is either the shrinkage or NOVELIST estimator, $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$-th largest eigenvalue and the corresponding eigenvector of the sample covariance matrix, respectively.

<!-- The number of principal components $K$ can be selected using information criteria. -->

Similar to the NOVELIST estimator, its PC-adjusted variant $\hat{\boldsymbol{W}}^{N, K}_{1}$ requires a cross-validation procedure to select the threshold paramter and adjustment to obtain positive definiteness. The reconciled forecasts for each cross-validation window are obtained using MinT with $\hat{\boldsymbol{W}}^{N, K}_{1}$.

@@

To address the potential issue of Problem [3](#prob3) of proportionality relationship $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$ not holding in practice. A simple relaxation retains the $1$-step-ahead correlation structure but allows the variances to scale differently with horizon. The scaled variance estimator is given by:

$$
\hat{\boldsymbol{W}}^{g, sv}_{h} = \boldsymbol{D}^{1/2}_h \hat{\boldsymbol{R}}^g_1 \boldsymbol{D}^{1/2}_h, \quad

\hat{\boldsymbol{R}}^g_1 = D^{-1/2}_1 \, g(\hat{\boldsymbol{W}}_1) \, D^{-1/2}_1,
$$

where $\boldsymbol{D}_h$ is a diagonal matrix comprising $\hat{\sigma}^2_{1,h} , \ldots, \hat{\sigma}^2_{n,h}$ on its diagonal, and $\hat{\sigma}^2_{i,h}$ is $h$-step-ahead base forecast error variance of the $i$-th series. $\hat{\boldsymbol{R}}^g_1$ is the estimated one-step correlation matrix using either shrinkage or NOVELIST estimator.

When NOVELIST is used to produce $h$-step-ahead reconciled forecasts, the cross-validation procedure is modified to evaluate the out-of-sample $h$-step-ahead reconciled forecasts for each window and estimate optimal threshold parameter (instead of evaluating $1$-step-ahead reconciled forecasts). The reconciled forecasts are obtained using MinT with $\hat{\boldsymbol{W}}^{N, sv}_{h}$.

@@

Another alternative is to directly estimate the covariance matrix from the $h$-step-ahead base forecast errors, without assuming any proportionality relationship:

$$
\hat{\boldsymbol{W}}^{g}_{h} = g(\hat{\boldsymbol{W}}_h),
$$

where $\hat{\boldsymbol{W}}_h$ is the covariance matrix of in-sample h-step-ahead base forecast errors, and $g(.)$ is either the shrinkage or NOVELIST estimator.

Similarly, if NOVELIST is used, the cross-validation procedure is modified to evaluate the out-of-sample $h$-step-ahead reconciled forecasts for each window. The reconciled forecasts are obtained using MinT with $\hat{\boldsymbol{W}}^{N}_{h}$.
